Urbanization affects web abundance and aggregation of a funnel weaver, *Agelenopsis pennsylvanica*

### 2022-11-28 manuscript/manuscript.Rmd
More work on the manuscript

```{r import data, eval=FALSE, include = FALSE}
webs <- read.table(file = "/Users/bjpessman/Documents/phd_research_code/Agelenopsis_aggregation/data/webs.txt", header = TRUE) 
webs <- webs %>% 
  mutate(ID = factor(ID),
         Location = factor(Location), 
         Location = fct_recode(Location, 
                               "Urban Forest" = "Forest", 
                               "Urban Center" = "Campus"),
         Location = fct_relevel(Location, "Urban Forest", "Urban Center"),
         Land = fct_recode(Land, 
                           "Urban, High" = "UrbanizedHighIntensity",
                           "Urban, Medium" = "UrbanizedMediumIntensity",
                           "Urban, Low" = "UrbanizedLowIntensity",
                           "Deciduous Forest" = "DeciduousForest",
                           "Woody Wetlands" = "WoodyWetlands"),
         Land = fct_relevel(Land, 
                            "Urban, High", 
                            "Urban, Medium", 
                            "Urban, Low", 
                            "Deciduous Forest", 
                            "Woody Wetlands"),
                  patch_area_km = patch_area_mm * 1e-6) %>% 
  dplyr::rename("Impervious" = "Imperv", 
                "Dist_to_Road" = "road_dist_data", 
                "Traffic_Dist_Road" = "dist_traffic_road", 
                "Dist_to_Highway" = "highway_dist_m", 
                "Traffic_Dist_Highway" = "dist_traffic_highway") #data on individual webs

sites <- webs %>% 
  filter(Web == "W001")

substrates <- read.table(file = "/Users/bjpessman/Documents/phd_research_code/Agelenopsis_aggregation/data/substrate.txt", header = TRUE) 
```

@@

Animal abundance and distribution are influenced by combinations of environmental factors, intra- and interspecific interactions, and dispersal capacity [@svamberkova_leps_2020; @diez_pulliam_2007]. Environmental factors -- including climatic conditions, land cover, habitat heterogeneity, and access to certain resources (like water and shelter) -- are suspected to relate to physiological constraints that limit the fundamental niche [@chase_leibold_2003]. Interactions within species (between conspecific competitors and mates) and across species (with predators, prey, and heterospecific competitors) further shape animal spatial patterns [@anderson_2017; @boulangeat_2012]. While dispersal capacity can determine a species’ range, barriers can limit dispersal and prevent animals from reaching suitable habitat  [@boulangeat_2012; @amundrud_2018]. Aggregations can form when suitable conditions and resources are not evenly distributed, leading to higher-quality patches [@churchill_2020]. For example, gaps in the tree canopy allow more sunlight to reach the understory -- thus, increasing plant height and soil moisture and decreasing leaf litter -- causing some invertebrate families to become more abundant in these patches [@perry_2018]. Clumping can also result from incomplete dispersal that occurs when barriers, often due to habitat structure, prevent adequate dispersal [@riechert_1974b]. The relative importance of different spatial predictors can vary by the type of taxon-specific spatial information being collected -- *i.e.* presence, abundance, aggregation, or vertical position [@boulangeat_2012].

Due to intense habitat disturbance and dense human presence, urbanization has drastically modified the land [@alberti_2020]; @mcdonnell_pickett_1990; @pickett_2001) which has altered environmental conditions, community dynamics, and dispersal capability as a result [@piano_2020]. As these factors determine animal spatial patterns, it is likely that urbanized landscapes have changed how animals spatially distribute themselves, in turn. Typically, city tree cover is reduced in areas where the potential natural vegetation is forested [@nowak_1996] and greater in areas with prairie, grassland, and desert origins [@mckinney_2006]. Urban environments are additionally characterized by decreased biodiversity [@mckinney_2002] and increased impervious cover [compacted soil from buildings and pavement; @nowak_greenfield_2020] and habitat patchiness [@cadenasso_2007]. These conditions can contribute to changes in climatic conditions, such as when impervious cover traps heat and increases the average temperatures in cities compared to surrounding areas –- *i.e.*, the Urban Heat Island Effect [@morabito_2021]. Disturbance from human presence can lead to air, water, and light pollution [@grimm_2008; @hopkins_2018], as well as air-borne, water-borne, and substrate-borne noise [@barber_2010; @slabbekoorn_2010; @wu_elias_2014]. There is growing evidence that urbanization changes species interactions, even reshaping whole food webs [@start_2020] and the landscape of parasite and disease transmission [@bradley_altizer_2007]. Further, habitat fragmentation and urban features often limit dispersal, which can cause higher degrees of aggregation for urban animals [@johnson_munshisouth_2017]. Urban ecosystems are extremely heterogeneous with unique and diverse habitats nested within [@rivkin_2019]. Habitats within urban ecosystems can vary considerably in feature, condition, and disturbance, resulting in variable access and distributions of resources that can differentially drive animal spatial patterns [@murray_2019; @trubl_2012].

Variation in the abundance and distribution of urban arthropods can have implications for ecosystem function due to the essential ecosystems services that arthropods provide, like pest control [@otoshi_2015] and pollination [@ivers_2022]. Arthropods are considered critical bioindicators of ecosystem health as their short generation time requires them to respond rapidly to human-induced change [@mcintyre_2000]. However, the impact of urbanization on arthropod ecology is still understudied [but see @kotze_2011; @fenoglio_2020], and the relationship between urban habitats and arthropod abundance and diversity remain controversial [@gippet_2017; @dale_frank_2018]. Such discrepancies are probably due to abundance and distributions being species- [@gippet_2017], habitat- [@bennett_lovell_2014], and scale-dependent [@philpott_2014]. While some animals benefit from urban conditions [@guenard_2015], many cannot survive the disturbance and change associated with cities [@wagner_2020]. Thus, it is important to determine how urban factors may be contributing to variation in arthropod abundance and spatial patterns for the conservation of arthropods and their essential services. Specifically, spiders play an important role in pest control and are predicted to collectively consume up to 800 million tons of arthropods each year [@nyffeler_birkhofer_2017]. Many web-building spiders are choosy in their web locations based on the distribution and abundance of other arthropod prey [@trubl_2012; @riechert_tracy_1975]. Territoriality is a form of intra-specific interactions that can also shape spatial distributions of web-builders [@riechert_1974b]. Further, previous studies have found environmental factors that vary across urban habitats may drive differences in spider spatial distribution, including noise [@bunkley_2017; @gomes_2020], pollution [@zmudzki_laskowski_2012], fragmentation [@stefani_delclaro_2015], microclimate conditions [@riechert_tracy_1975], vegetation structure [@peng_2020], and suitable web substrates [@chan_2009].

The mechanisms behind solitary spider aggregations are unclear despite reports of aggregations of spiders in urban habitats, such as web-builders *Lactrodectus hesperus* and *Brigittea civica*. This gap in knowledge is concerning as urban spider aggregations can have medical and aesthetic consequences, and implications for urban pest control services. Prey abundance seems to be an important driver because it has been shown to be directly and indirectly related to urban spider aggregations, the latter of which is shown through investigations as light sources at night as cues of prey aggregations. Random spacing in solitary spiders is typically maintained by territoriality and aggression over resources like prey and mates. Occassionally, conditions arise that allow solitary spiders to become more tolerant of neighbors, and understanding these contexts can help uncover the mechanisms of aggregative behavior. Aggregations have been shown to form as a result of dispersal limitations, predator release, lekking reproductive strategies, plant structure, and environmental stressors. The degree of aggregation in spiders can also change over time, often times more intense during juvenile stages when dispersal is limited, spreading over time as subadults and adults become more territorial []. Studies on mechanisms of aggregations in spiders are presented and discussed in Table 1.

Funnel weavers in the genus *Agelenopsis* actively select web locations based on experience of perceived quality of the site [@riechert_1976], and are thus likely impacted environmental conditions resulting from urbanization. Prey abundance, available web substrates, vegetation structure, and microclimate conditions have been shown to be indicators of a quality web site in a desert habitat for *A. aperta* [@riechert_1974; @riechert_1973]. In the same species, spatial patterns changed throughout life stages, although this may depend on environmental conditions and dispersal [@riechert_1974]. *Agelenopsis pennsylvanica* build webs that consist of a sheet that functions as a prey capture area and a retreat to escape from predators and unfavorable conditions (Fig. 1), like direct sunlight [@riechert_tracy_1975]. The horizontal sheet is not sticky, and the web is used solely for its vibratory properties to rapidly and accurately locate prey that fall upon or walk across the web [@singer_2000]. Occasionally, vertical silk strands are found that may function to capture flying prey or evade predators [@singer_2000]. The spider sits in the retreat until the prey is sensed and returns to the retreat immediately after capturing the prey. Juvenile males and females build semi-permanent webs, but males leave their web shortly after maturity in search of a receptive female [@singer_2000].

@@

We conducted our study in September and October of 2020 in Lincoln, Nebraska, USA. Lincoln -- with a population of about 300,000 people [@census] -- has diverse urban habitats that vary in the degree of disturbance, management, and plant cover. We compared the abundance, aggregation, and position of funnel webs (*Agelenopsis pennsylvanica*) between two urban habitats: the University of Nebraska-Lincoln (UNL) City Campus (an urban center) and Wilderness Park (an urban forest) (Fig. 2, Fig. S1). These two habitats were selected due to their distinctive characteristics (described below), preliminary observations of potential differences in funnel weaver abundance and aggregation, and ease of access.

To first quantify differences in the environmental conditions between urban habitats, we gathered and analyzed Geographic Information Systems (GIS) data in QGIS (v. 3.16.3-Hannover, ESRI 102704). We obtained tree cover [@usfs_2016], impervious (building and pavement) cover [@usgs_2019_imperv], and land use class [@usgs_2019_land] from the National Land Cover Database (30-meter resolution). We used an image from path 28, row 32 of the Landsat-7 World Reference System-2 (Band 6) to collect spectral radiance -- a measure of thermal conditions on Earth's surface (30-meter resolution, taken on September 2, 2020). Finally, we assessed artificial light at night levels using the 2020 Visible Infrared Imaging Radiometer Suite (VIIRS) from https://www.lightpollutionmap.info/ (500-meter resolution). We polygonized the tiff files and calculated the average value for each factor by habitat: UNL City Campus, Wilderness Park, and Lincoln City Limits (see Fig. 2, Fig. S2-S5). We included measurements for Lincoln city limits to show how the two study habitats compare to the city of Lincoln overall. We calculated the averages using the following equation:

@@

We defined the study boundaries of UNL City Campus as Salt Creek Roadway, North Antelope Valley Parkway, North 10th Street, and Q Street (1.12 $km^2$, Fig. 2, Fig. S1). UNL City Campus is adjacent to, and just north of, downtown Lincoln (253.73 $km^2$). Impervious cover -- in the form of buildings and pavement -- dominates the land on UNL City Campus, considerably more than impervious cover across the city of Lincoln (Table 1, Fig. S2). Lawns are maintained by frequent mowing, and gardens and landscaping are regularly pruned. Tree cover on UNL City Campus and in Lincoln overall is low (< 6%, Table 1, Fig. S3). There was heavy foot traffic from UNL's 20,286 undergraduates enrolled in the 2020-2021 academic year, 37% of which lived in college-owned, -operated, or -affiliated housing [@iea_2021]. The entire area of UNL City Campus was categorized as urban with varying intensity (Table 1, Fig. 2). Over half of UNL City Campus was high intensity urban cover compared to only 13.16% for Lincoln overall. Further, UNL City Campus exhibits heighten spectral radiance (thermal conditions, Fig. S4) and artificial light at night radiance (Fig. S5), even above that of Lincoln as a whole (Table 1). Thus, we designated UNL City Campus as an *urban center* and refer to it as such hereafter.

Wilderness Park is a stretch of forested land (6.44 $km^2$, Fig. 2, Fig. S1) starting about 3.58 km southwest of UNL City Campus and downtown Lincoln, extending south and along Lincoln city limits. We found about two-fifths of Wilderness Park to be shrouded in trees (Table 1, Fig. S3), with less than 2% of its land imperviously covered (Table 1, Fig. S2). Over three quarters of Wilderness Park is categorized as forested land with only 5.44% labeled as urban (Table 1, Fig. 2). Wilderness Park has about 18 kilometers of dirt trails, dense deciduous tree growth, and variable undergrowth. The trails are open for recreational walking, running, biking, and horse riding. While we found no public records of the number of visitors to Wilderness Park, trail-goers were infrequent during data collection. Wilderness Park showed greatly reduced spectral radiance (Fig. S4) and artificial light at night radiance (Fig. S5) than the city of Lincoln (Table 1). Due to its proximity to Lincoln city limits, we designate Wilderness Park as an *urban forest* and refer to it as such hereafter.

@@

Within the boundaries of the urban forest, we assigned numbers to intersections of three or more trails (Fig. S6). Similarly, we assigned numbers to each intersection of roads and sidewalks that meet from the four cardinal directions in the urban center (Fig. S6).  We used a random number generator to randomly select 10 intersections in the urban forest and 12 intersections in the urban center as start sites (Table S1). We marked the GPS coordinates of these locations using a GPS receiver (Magellan SporTrak Map, Model 980616-01, Parsippany, New Jersey, USA).

From each start site, we began walking along a path in a randomly selected direction (North, East, South, or West using a random number generator). We thoroughly searched the edges of the path for funnel webs as the GPS receiver tracked our movement (Fig. S7). We walked along pre-existing sidewalks, roads, and trails because paths were a common feature of both habitats along which we had previously found funnel weavers. We often find *A. pennsylvanica* in edge habitats in the study region, and this tendency to build webs at a habitat's edge is shared by other spider species [@cobbold_supp_2012]. Additionally, the thick undergrowth of the urban forest often restricted searches off the trails. To ensure equivalent search efforts, we took more time searching the edge of dense undergrowth in the urban forest.

@@

Search distance can often be used to estimate abundance [@buckland_2005], as spiders are likely less abundant in an area where we need to search further from the start site to the focal web. Thus, we use search distance as evidence for variation in abundance. To quantitatively determine the abundance of webs and spiders near each focal web, we measured a ten-meter radius around the focal web and counted the number of occupied and unoccupied webs determined in the same manner as the focal web. We also assessed nearest neighbor distances as a measure of aggregation behavior. Starting at the focal web and working outwards, we marked the locations of the two webs (occupied and unoccupied) nearest to the focal web within ten meters. We limited our measurements to the nearest two webs because we often found no more than three webs (including the focal web) per site in the urban forest. We measured the shortest distance from the retreat of the focal web to the retreat of the two nearest neighbor webs. To explore differences in web placement, we measured the web height from the ground to the web retreat. Differences in web height have been shown to vary by environmental factors associated with urbanization [@dahirel_2017]. To get an idea of plant species richness at each site, we recorded each unique plant species found in a ten-meter radius of the focal web (Fig. S8).

We collected data from September 23 to October 14, 2020. During these dates, all *A. pennsylvanica* have reached maturity [@guarisco_2014; unpublished data]. Of the 64 occupied webs, 61 were occupied by females indicating that males had likely abandoned their webs in search of females. The 67 unoccupied webs could be those once owned by males or those abandoned by females. While webs were found on many of the same plant types between sites in the urban forest and urban center in varying proportions, 15% of webs in urban center sites were built on non-plant substrates (Fig. S9).

@@

To assess potential mechanisms behind variation in abundance, aggregation, and position, we looked for correlations between spider spatial patterns and environmental conditions -- collected in the field (plant species richness) or using GIS for each site. In QGIS, we created 100 meter buffers around focal web sites and processed the GIS variables mentioned above (tree cover, impervious cover, spectral radiance, and artificial light at night radiance) in the same manner as before. To these we added patch area, potential disturbance from nearby roads and highways, and road length in the 100 meter buffer. We used a 100 meter buffer to accommodate the resolution of the GIS data. We refrained from including a larger buffer region as the dispersal range of this system is unknown and may be dependent on barriers present in the habitat.

@@

For the first set of models, we focused on whether web abundance, aggregation, and position differ between the two habitats (the urban center and urban forest). We used Poisson generalized linear models [*stats* package in R, @stats] for the response variables: search distance, number of webs, and number of spiders. We used Poisson generalized linear mixed effect models [*lme4* package in R, @lme4] with site as a random effect for the response variables: nearest neighbor distance and web height. For the model with nearest neighbor distance, we also included neighbor number (nearest and second nearest web) and its interaction with habitat. We used Poisson distributions because the response variables were not normally distributed according to Shapiro-Wilk Normality Tests, and all data were strictly positive with no upper limit.

@@

Since we found significant variation in distribution measures between habitats, we wanted to explore potential environmental mechanisms of such variation within habitats. We first tested whether the nine environmental predictors varied by habitat (Table 2). Each predictor was tested by habitat using Poisson generalized linear models as the predictors were positive and not normally distributed. Each environmental predictor significantly varied by habitat (Table 2), so we subset the data by habitat. Among the environmental predictors, we removed variables that were highly correlated (cutoff of 0.6) using the *caret* package in R [@caret]. Artificial light at night radiance was dropped from all analyses for being highly correlated. Only plant species richness, road disturbance, and highway disturbance remained for models using all data. We kept tree cover, plant species richness, patch area, and road and highway disturbance for the urban forest subset. The urban center subset includes plant species richness, patch area, spectral radiance, road and highway disturbance, and road length.

@@

We used the habitat subsets and the selected environmental predictor variables to assess correlations between predictor variables and spatial pattern measures. In the global models, we included the predictors that remained after removal for collinearity (no interactions) with the response variables: web and spider abundance, nearest neighbor distance, and web height for the complete data and habitat subsets. We assessed nearest neighbor distance and excluded second nearest neighbor distance because both were significantly different between habitats. We excluded search distance as a response variable because the predictors were collected from the site of the focal web and may vary along the search path. Several predictors (plant species richness, patch area, and road and highway disturbance) were highly skewed toward zero, so we used a natural log-transformation prior to model selection. We used Poisson generalized linear models for number of webs and spiders and nearest neighbor distance and Poisson generalized linear mixed-effect models for web height with site as a random effect. From these global models, we performed Akaike Information Criterion (AIC) model selection with the *MuMIn* package in R [@MuMIn], which ranks models produced from every combination of the predictors in the global model. We report here the results of each top model, as well as a model average of models within 2 delta AICc points of the top model. Reported estimates and standard error are back-transformed for variables that were natural log-transformed using the *exp()* function. We checked that the Variance of Inflation (VIF) stayed below a score of two for predictors in each top model using the *car* package in R [@car].

We also explore how land use may affect web abundance, aggregation, and position using the National Land Cover Database and QGIS to record the category of the land use type for the pixel that each site was located. We tested Poisson generalized linear models for each response variable with land cover as the only predictor. We include these results in supplemental material (Fig. S11, Table S3-7) as the sample sizes were small (deciduous forest = 7, woody wetlands = 3, urban high intensity = 4, urban medium intensity = 6, urban low intensity = 2).

>>>>

@@

Animal abundance and distribution are influenced by combinations of environmental factors, intra- and interspecific interactions, and dispersal capacity [@svamberkova_leps_2020; @diez_pulliam_2007]. Environmental factors -- including climatic conditions, land cover, habitat heterogeneity, and access to certain resources (like water and shelter) -- are suspected to relate to physiological constraints that limit the fundamental niche [@chase_leibold_2003]. Interactions within species (between conspecific competitors and mates) and across species (with predators, prey, and heterospecific competitors) further shape animal spatial patterns [@anderson_2017; @boulangeat_2012]. While dispersal capacity can determine a species’ range, barriers can limit dispersal and prevent animals from reaching suitable habitat  [@boulangeat_2012; @amundrud_2018].

Aggregations can form when suitable conditions and resources are not evenly distributed, leading to higher-quality patches [@churchill_2020]. For example, gaps in the tree canopy allow more sunlight to reach the understory -- thus, increasing plant height and soil moisture and decreasing leaf litter -- causing some invertebrate families to become more abundant in these patches [@perry_2018]. Clumping can also result from incomplete dispersal that occurs when barriers, often due to habitat structure, prevent adequate dispersal [@riechert_1974b]. The relative importance of different spatial predictors can vary by the type of taxon-specific spatial information being collected -- *i.e.* presence, abundance, aggregation, or vertical position [@boulangeat_2012].

Due to intense habitat disturbance and dense human presence, urbanization has drastically modified the land [@alberti_2020]; @mcdonnell_pickett_1990; @pickett_2001) which has altered environmental conditions, community dynamics, and dispersal capability as a result [@piano_2020]. As these factors determine animal spatial patterns, it is likely that urbanized landscapes have changed how animals spatially distribute themselves, in turn. Typically, city tree cover is reduced in areas where the potential natural vegetation is forested [@nowak_1996] and greater in areas with prairie, grassland, and desert origins [@mckinney_2006]. Urban environments are additionally characterized by decreased biodiversity [@mckinney_2002] and increased impervious cover [compacted soil from buildings and pavement; @nowak_greenfield_2020] and habitat patchiness [@cadenasso_2007]. These conditions can contribute to changes in climatic conditions, such as when impervious cover traps heat and increases the average temperatures in cities compared to surrounding areas –- *i.e.*, the Urban Heat Island Effect [@morabito_2021].

Disturbance from human presence can also lead to air, water, and light pollution [@grimm_2008; @hopkins_2018], as well as air-borne, water-borne, and substrate-borne noise [@barber_2010; @slabbekoorn_2010; @wu_elias_2014]. There is growing evidence that urbanization changes species interactions, even reshaping whole food webs [@start_2020] and the landscape of parasite and disease transmission [@bradley_altizer_2007]. Further, habitat fragmentation and urban features often limit dispersal, which can cause higher degrees of aggregation for urban animals [@johnson_munshisouth_2017]. Urban ecosystems are extremely heterogeneous with unique and diverse habitats nested within [@rivkin_2019]. Habitats within urban ecosystems can vary considerably in feature, condition, and disturbance, resulting in variable access and distributions of resources that can differentially drive animal spatial patterns [@murray_2019; @trubl_2012].

Variation in the abundance and distribution of urban arthropods can have implications for ecosystem function due to the essential ecosystems services that arthropods provide, like pest control [@otoshi_2015] and pollination [@ivers_2022]. Arthropods are considered critical bioindicators of ecosystem health as their short generation time requires them to respond rapidly to human-induced change [@mcintyre_2000]. However, the impact of urbanization on arthropod ecology is still understudied [but see @kotze_2011; @fenoglio_2020], and the relationship between urban habitats and arthropod abundance and diversity remain controversial [@gippet_2017; @dale_frank_2018]. Such discrepancies are probably due to abundance and distributions being species- [@gippet_2017], habitat- [@bennett_lovell_2014], and scale-dependent [@philpott_2014]. While some animals benefit from urban conditions [@guenard_2015], many cannot survive the disturbance and change associated with cities [@wagner_2020]. Thus, it is important to determine how urban factors may be contributing to variation in arthropod abundance and spatial patterns for the conservation of arthropods and their essential services.

Specifically, spiders play an important role in pest control and are predicted to collectively consume up to 800 million tons of arthropods each year [@nyffeler_birkhofer_2017]. Many web-building spiders are choosy in their web locations based on the distribution and abundance of other arthropod prey [@trubl_2012; @riechert_tracy_1975]. Territoriality is a form of intra-specific interactions that can also shape spatial distributions of web-builders [@riechert_1974b]. Further, previous studies have found environmental factors that vary across urban habitats may drive differences in spider spatial distribution, including noise [@bunkley_2017; @gomes_2020], pollution [@zmudzki_laskowski_2012], fragmentation [@stefani_delclaro_2015], microclimate conditions [@riechert_tracy_1975], vegetation structure [@peng_2020], and suitable web substrates [@chan_2009].

**This paragraph needs citations, but it will be easier after finishing the aggregation table**
The mechanisms behind solitary spider aggregations are unclear despite reports of aggregations of spiders in urban habitats, such as web-builders *Lactrodectus hesperus* and *Brigittea civica* [@johnson_et_al_2012; @mammola_et_al_2018]. This gap in knowledge is concerning as urban spider aggregations can have medical and aesthetic consequences, and implications for urban pest control services. Prey abundance seems to be an important driver because it has been shown to be directly [] and indirectly [] related to urban spider aggregations, the latter of which is shown through investigations as light sources at night as cues of prey aggregations []. Random spacing in solitary spiders is typically maintained by territoriality and aggression over resources like prey and mates [@riechert_1974b]. Occassionally, conditions arise that allow solitary spiders to become more tolerant of neighbors, and understanding these contexts can help uncover the mechanisms of aggregative behavior. Aggregations have been shown to form as a result of dispersal limitations [], predator release [], lekking reproductive strategies [], plant structure [], and environmental stressors []. The degree of aggregation in spiders can also change over time, often times more intense during juvenile stages when dispersal is limited, spreading over time as subadults and adults become more territorial [@riechert_1974b]. Studies on mechanisms of aggregations in spiders are presented and discussed in Table 1.

Funnel weavers in the genus *Agelenopsis* actively select web locations based on experience of perceived quality of the site [@riechert_1976], and are thus likely impacted by environmental conditions resulting from urbanization. Prey abundance, available web substrates, vegetation structure, and microclimate conditions have been shown to be indicators of a quality web site in a desert habitat for *A. aperta* [@riechert_1974; @riechert_1973]. *Agelenopsis pennsylvanica* build webs that consist of a sheet that functions as a prey capture area and a retreat to escape from predators and unfavorable conditions (Fig. 1), like direct sunlight [@riechert_tracy_1975]. The horizontal sheet is not sticky, and the web is used solely for its vibratory properties to rapidly and accurately locate prey that fall upon or walk across the web [@singer_2000]. Occasionally, vertical silk strands are found that may function to capture flying prey or evade predators [@singer_2000]. The spider sits in the retreat until the prey is sensed and returns to the retreat immediately after capturing the prey. Juvenile males and females build semi-permanent webs, but males leave their web shortly after maturity in search of a receptive female [@singer_2000].

@@

We conducted our study in September and October of 2020 in Lincoln, Nebraska, USA. Lincoln -- with a population of about 300,000 people [@census] -- has diverse urban habitats that vary in the degree of disturbance, management, and plant cover. We compared the abundance, aggregation, and position of funnel webs (*Agelenopsis pennsylvanica*) between two urban habitats: the University of Nebraska-Lincoln (UNL) City Campus (an urban center) and Wilderness Park (an urban forest) (Fig. 2, Fig. S1). These two habitats were selected due to their distinctive characteristics (described below), preliminary observations of potential differences in funnel weaver abundance and aggregation (Fig. S2), and ease of access.

To first quantify differences in the environmental conditions between urban habitats, we gathered and analyzed Geographic Information Systems (GIS) data in QGIS (v. 3.16.3-Hannover, ESRI 102704). We obtained tree cover [@usfs_2016], impervious (building and pavement) cover [@usgs_2019_imperv], and land use class [@usgs_2019_land] from the National Land Cover Database (30-meter resolution). We used an image from path 28, row 32 of the Landsat-7 World Reference System-2 (Band 6) to collect spectral radiance -- a measure of thermal conditions on Earth's surface (30-meter resolution, taken on September 2, 2020). Finally, we assessed artificial light at night levels using the 2020 Visible Infrared Imaging Radiometer Suite (VIIRS) from https://www.lightpollutionmap.info/ (500-meter resolution). We polygonized the tiff files and calculated the average value for each factor by location: UNL City Campus, Wilderness Park, and Lincoln City Limits (see Fig. 2, Fig. S3-S6). We included measurements for Lincoln city limits to show how the two study locations compare to the city of Lincoln overall. We calculated the averages using the following equation:

@@

We defined the study boundaries of UNL City Campus as Salt Creek Roadway, North Antelope Valley Parkway, North 10th Street, and Q Street (1.12 $km^2$, Fig. 2, Fig. S1). UNL City Campus is adjacent to, and just north of, downtown Lincoln (253.73 $km^2$). Impervious cover -- in the form of buildings and pavement -- dominates the land on UNL City Campus, considerably more than impervious cover across the city of Lincoln (Table 1, Fig. S3). Lawns are maintained by frequent mowing, and gardens and landscaping are regularly pruned. Tree cover on UNL City Campus and in Lincoln overall is low (< 6%, Table 1, Fig. S4). There was heavy foot traffic from UNL's 20,286 undergraduates enrolled in the 2020-2021 academic year, 37% of which lived in college-owned, -operated, or -affiliated housing [@iea_2021]. The entire area of UNL City Campus was categorized as urban with varying intensity (Table 1, Fig. 2). Over half of UNL City Campus was high intensity urban cover compared to only 13.16% for Lincoln overall. Further, UNL City Campus exhibits heighten spectral radiance (thermal conditions, Fig. S5) and artificial light at night radiance (Fig. S6), even above that of Lincoln as a whole (Table 1). Thus, we designated UNL City Campus as an *urban center* and refer to it as such hereafter.

Wilderness Park is a stretch of forested land (6.44 $km^2$, Fig. 2, Fig. S1) starting about 3.58 km southwest of UNL City Campus and downtown Lincoln, extending south and along Lincoln city limits. We found about two-fifths of Wilderness Park to be shrouded in trees (Table 1, Fig. S4), with less than 2% of its land imperviously covered (Table 1, Fig. S2). Over three quarters of Wilderness Park is categorized as forested land with only 5.44% labeled as urban (Table 1, Fig. 2). Wilderness Park has about 18 kilometers of dirt trails, dense deciduous tree growth, and variable undergrowth. The trails are open for recreational walking, running, biking, and horse riding. While we found no public records of the number of visitors to Wilderness Park, trail-goers were infrequent during data collection. Wilderness Park showed greatly reduced spectral radiance (Fig. S5) and artificial light at night radiance (Fig. S6) than the city of Lincoln (Table 1). Due to its proximity to Lincoln city limits, we designate Wilderness Park as an *urban forest* and refer to it as such hereafter.

@@

Within the boundaries of the urban forest, we assigned numbers to intersections of three or more trails (Fig. S7). Similarly, we assigned numbers to each intersection of roads and sidewalks that meet from the four cardinal directions in the urban center (Fig. S7).  We used a random number generator to randomly select 10 intersections in the urban forest and 12 intersections in the urban center as start sites (Table S1). We marked the GPS coordinates of these locations using a GPS receiver (Magellan SporTrak Map, Model 980616-01, Parsippany, New Jersey, USA).

From each start site, we began walking along a path in a randomly selected direction (North, East, South, or West using a random number generator). We thoroughly searched the edges of the path for funnel webs as the GPS receiver tracked our movement (Fig. S8). We walked along pre-existing sidewalks, roads, and trails because paths were a common feature of both habitats along which we had previously found funnel weavers. We often find *A. pennsylvanica* in edge habitats in the study region, and this tendency to build webs at a habitat's edge is shared by other spider species [@cobbold_supp_2012]. Additionally, the thick undergrowth of the urban forest often restricted searches off the trails. To ensure equivalent search efforts, we took more time searching the edge of dense undergrowth in the urban forest.

@@

Search distance can often be used to estimate abundance [@buckland_2005], as spiders are likely less abundant in an area where we need to search further from the start site to the focal web. Thus, we use search distance as evidence for variation in abundance. To quantitatively determine the abundance of webs and spiders near each focal web, we measured a ten-meter radius around the focal web and counted the number of occupied and unoccupied webs determined in the same manner as the focal web. We also assessed nearest neighbor distances as a measure of aggregation behavior. Starting at the focal web and working outwards, we marked the locations of the two webs (including occupied and unoccupied) nearest to the focal web within ten meters. We limited our measurements to the nearest two webs because we often found no more than three webs (including the focal web) per site in the urban forest. We measured the shortest distance from the retreat of the focal web to the retreat of the two nearest neighbor webs. To explore differences in web placement, we measured the web height from the ground to the web retreat for every web with the ten-meter radius. Differences in web height have been shown to vary by environmental factors associated with urbanization [@dahirel_2017]. To get an idea of plant species richness at each site, we recorded each unique plant species found in a ten-meter radius of the focal web (Fig. S8).

We collected data from September 23 to October 14, 2020. During these dates, all *A. pennsylvanica* have reached maturity [@guarisco_2014; personal observation]. Of the 64 occupied webs, 61 were occupied by females indicating that males had likely abandoned their webs in search of females. The 67 unoccupied webs could be those once owned by males or those abandoned by females. While webs were found on many of the same plant types between sites in the urban forest and urban center in varying proportions, 15% of webs in urban center sites were built on non-plant substrates (Fig. S9).

@@

To assess potential mechanisms behind variation in abundance, aggregation, and position, we looked for correlations between spider spatial patterns and environmental conditions -- collected in the field (plant species richness) or using GIS for each site. In QGIS, we created 100 meter buffers around each focal web location and processed the GIS variables mentioned above (tree cover, impervious cover, spectral radiance, and artificial light at night radiance) in the same manner as before. To these we added patch area, potential disturbance from nearby roads and highways, and road length in the 100 meter buffer. We used a 100 meter buffer to accommodate the resolution of the GIS data. We refrained from including a larger buffer region as the dispersal range of this system is unknown and may be dependent on barriers present in the habitat.

@@

For the first set of models, we focused on whether web abundance, aggregation, and position differ between the two habitats (the urban center and urban forest). We used Poisson generalized linear models [*stats* package in R, @stats] for the response variables: search distance, number of webs, and number of spiders. We used Poisson generalized linear mixed effect models [*lme4* package in R, @lme4] with site as a random effect for the response variables: nearest neighbor distance and web height. For the model with nearest neighbor distance, we also included neighbor number (nearest and second nearest web) and its interaction with habitat We used Poisson distributions because the response variables were not normally distributed according to Shapiro-Wilk Normality Tests, and all data were strictly positive with no upper limit.

@@

Since we found significant variation in distribution measures between habitats, we wanted to explore potential environmental mechanisms of such variation within the test habitats. We first tested whether the nine environmental predictors varied by habitat (Table 2). Each predictor was tested by habitat using Poisson generalized linear models as the predictors were positive and not normally distributed. Each environmental predictor significantly varied by habitat (Table 2), so we subset the data by habitat. Among the environmental predictors, we removed variables that were highly correlated (cutoff of 0.6) using the *caret* package in R [@caret]. Artificial light at night radiance was dropped from all analyses for being highly correlated. Only plant species richness, road disturbance, and highway disturbance remained for models using all data. We kept tree cover, plant species richness, patch area, and road and highway disturbance for the urban forest subset. The urban center subset includes plant species richness, patch area, spectral radiance, road and highway disturbance, and road length.

@@

We used the habitat subsets and the selected environmental predictor variables to assess correlations between predictor variables and spatial pattern measures. In the global models, we included the predictors that remained after removal for collinearity (no interactions) with the response variables: web and spider abundance, nearest neighbor distance, and web height for the complete data and habitat subsets. We assessed nearest neighbor distance and excluded second nearest neighbor distance because both were significantly different between habitats. We excluded search distance as a response variable because the predictors were collected from the site of the focal web and may vary along the search path. Several predictors (plant species richness, patch area, and road and highway disturbance) were highly right-skewed, so we used a natural log-transformation prior to model selection. We used Poisson generalized linear models for number of webs and spiders and nearest neighbor distance and Poisson generalized linear mixed-effect models for web height with site as a random effect. From these global models, we performed Akaike Information Criterion (AIC) model selection with the *MuMIn* package in R [@MuMIn], which ranks models produced from every combination of the predictors in the global model. We report here the results of each top model, as well as a model average of models within 2 delta AICc points of the top model. We reported the estimates and standard error from the models without back-transformations for the log-transformed variables. We checked that the Variance of Inflation (VIF) stayed at or below a score of two for predictors in each top model using the *car* package in R [@car].

We also explore how land use may affect web abundance, aggregation, and position using the National Land Cover Database and QGIS to record the category of the land use type for the pixel that each site was located. We tested Poisson generalized linear models for each response variable with land cover as the only predictor. We used the *Anova()* function in the *car* package [@car] to get the Likelihood Ratio Chi-Squared values, then used the *glht()* funtion in the *multcomp* package [@multcomp] to perform pairwise comparisons among categories. We include these results in supplemental material (Fig. S12, Table S3) as the sample sizes were small (deciduous forest = 7, woody wetlands = 3, urban high intensity = 4, urban medium intensity = 6, urban low intensity = 2).
